---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
author: "Nerea Martinez Yarza"
date: "2023-10-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Code for reading the dataset and processing the data
```{r cars}
data <- read.csv("activity.csv")
summary(data)
```
## Question 1 - What is mean total number of steps taken per day?
```{r}
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```

#### Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(total.steps, binwidth=1000, xlab="total steps per day")
```

#### Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## Question 2 - What is the average daily activity pattern?
#### Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
average <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=average, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps")
```

## Question 3 - Imputing missing values
#### Calculate and report the total number of missing values in the dataset
```{r}
missingValues <- is.na(data$steps)
table(missingValues)
```

#### Devise a strategy for filling in all of the missing values in the dataset
In this case we replace each missing value with the mean value of its 5-minute interval
```{r}
fill.value <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (average[average$interval==interval, "steps"])
  return(filled)
}
```

#### Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r} 
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

#### Make a histogram of the total number of steps per day and calculate the mean and median total number of steps taken per day
```{r}
stepsTotal <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(stepsTotal, binwidth=1000, xlab="total number of steps per day")
mean(stepsTotal)
median(stepsTotal)
```

## Question 4 - Are there differences in activity patterns between weekdays and weekends?
#### Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
```{r}
filled.data$dateType <-  ifelse(as.POSIXlt(filled.data$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

#### Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken
```{r}
filled.dataAverage <- aggregate(steps ~ interval + dateType, data=filled.data, mean)
ggplot(filled.dataAverage, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("avarage steps")
```
